---
title: "Monitor Lizard Diversification with DAISIE"
author: "Ian G. Brennan & Carlos Pavon Vazquez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  #html_document:
  #  toc: yes
  pdf_document:
    toc: yes
linkcolor: "magenta"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(root.dir = "~/Documents/GitHub/MonitorDiversification/Data")
```

```{r wrap-hook, eval=T, echo=F}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

# Read Me
This is a living markdown document that will hopefully walk us through the data and code necessary to repeat the analyses of monitor lizard diversification using DAISIE. All the files and code are available at the GitHub Repository [*MonitorDiversification*](https://github.com/IanGBrennan/MonitorDiversification) (this is currently set to **private**).

\pagebreak

Start off by loading a few packages that we'll need along the way.
```{r, message=F, warning=F}
library(dplyr)
library(treeplyr)
library(RCurl)
library(phytools)
library(RColorBrewer)
library(ggplot2)
library(DAISIE)
# remember 'plyr' and 'dplyr' conflict, so don't load 'plyr'
```

```{r, include=FALSE, echo=FALSE}
# CHANGE THIS LOCATION IF YOU'RE RUNNING THE CODE ON YOUR OWN COMPUTER!
require("knitr")
opts_knit$set(root.dir = "~/Documents/GitHub/MonitorDiversification/Data")
```


# Data Visualization

Start by reading in the tree and visualize the distribution of taxa
```{r, eval=T, fig.height=10}
vtree <- read.tree("UsedTree.tre");
regions <- read.delim("geo_file_Regions.txt", sep="\t", row.names = 1)
dotTree(vtree, regions, labels=T,length=8)
```

\  

Read in the morphological data we have and visualize data.  
I've made an additional data column to look at the Australian monitor data. I've categorized them as "endemic" to Australia (most of the *Odatria/Varanus* subgenera), "immigrant" (*doreanus/chlorostigma/keithhornei*), or "extralimital" which is just all things existing elsewhere exclusively.
```{r, eval=T}
morph.data <- read.csv("MorphForIan.csv", header=T)
```

```{r, eval=T}
ggplot(morph.data, aes(Mean_logSVL, PC1_LinearMeasurements)) +
  geom_point(data = transform(morph.data[c("Mean_logSVL", 
                                           "PC1_LinearMeasurements")]), 
             color = "grey85") +
  geom_point(color = "#F46D43", size = 2) +
  #stat_ellipse() +
  theme_bw() +
  facet_wrap(vars(Oz_Group))
```

\pagebreak

```{r, eval=T}
ggplot(morph.data, aes(PC1_LinearMeasurements, PC1_DorsalViewHead)) +
  geom_point(data = transform(morph.data[c("PC1_LinearMeasurements", 
                                           "PC1_DorsalViewHead")]), 
             color = "grey85") +
  geom_point(color = "#66C2A5", size = 2) +
  #stat_ellipse() +
  theme_bw() +
  facet_wrap(vars(Oz_Group))
```

\pagebreak

```{r, eval=T}
ggplot(morph.data, aes(PC1_DorsalViewHead, PC1_LateralViewHead)) +
  geom_point(data = transform(morph.data[c("PC1_DorsalViewHead", 
                                           "PC1_LateralViewHead")]), 
             color = "grey85") +
  geom_point(color = "#3288BD", size = 2) +
  #stat_ellipse() +
  theme_bw() +
  facet_wrap(vars(Oz_Group))
```

\pagebreak

```{r, eval=T}
ggplot(morph.data, aes(PC1_LinearMeasurements, PC1_LateralViewHead)) +
  geom_point(data = transform(morph.data[c("PC1_LinearMeasurements", 
                                           "PC1_LateralViewHead")]), 
             color = "grey85") +
  geom_point(color = "#D53E4F", size = 2) +
  #stat_ellipse() +
  theme_bw() +
  facet_wrap(vars(Oz_Group))
```

\  

### Quick Take-away Thought:
From a preliminary view, it doesn't look like the three immigrant lineages (*doreanus, chlorostigma, keithhornei*) are exceptional in any of the morphological dimensions. Maybe that makes sense as to why they haven't speciated in Australia? They really just make it in around the edges, and they aren't doing anything wildly novel, so they don't spread/speciate?

\pagebreak

# Data Preparation
## Extracting branching times
We need to create an input file for DAISIE, consisting of the ages/names of incumbent clades and immigrant groups. 

If you have questions about implementing the DAISIE models, we can check with the package tutorial:
```{r, eval=F}
DAISIE_tutorial()
```

The radiation of subgenera *Varanus/Odatria* is the incumbent group, and there are three immigrant lineages (*V. keithhornei*, *V. doreanus*, *V. chlorostigma*).  
We want to get the ages of these lineages in the Australian region.
```{r, eval=T}
ov.bt <- branching.times(extract.clade(vtree, 
                                       getMRCA(vtree, c("komodoensis", "tristis"))))
names(ov.bt) <- NULL; ov.bt <- sort(ov.bt, decreasing=T)

vk.bt <- max(nodeHeights(vtree)) - 
         nodeheight(vtree, getParent(vtree, which(vtree$tip.label == "keithhornei")))
vd.bt <- max(nodeHeights(vtree)) - 
         nodeheight(vtree, getParent(vtree, which(vtree$tip.label == "doreanus")))
vc.bt <- max(nodeHeights(vtree)) - 
         nodeheight(vtree, getParent(vtree, which(vtree$tip.label == "chlorostigma")))
```

Let's do the same thing for the Papuan radiation
```{r, eval=T}
pa.bt <- branching.times(extract.clade(vtree, 
                                       getMRCA(vtree, c("indicus", "prasinus"))))
names(pa.bt) <- NULL; pa.bt <- sort(pa.bt, decreasing=T)

vsa.bt <- max(nodeHeights(vtree)) - 
         nodeheight(vtree, getParent(vtree, which(vtree$tip.label == "salvadorii")))
vsc.bt <- max(nodeHeights(vtree)) - 
         nodeheight(vtree, getParent(vtree, which(vtree$tip.label == "scalaris")))
vst.bt <- max(nodeHeights(vtree)) - 
         nodeheight(vtree, getParent(vtree, which(vtree$tip.label == "salvator")))
vpa.bt <- max(nodeHeights(vtree)) - 
         nodeheight(vtree, getParent(vtree, which(vtree$tip.label == "panoptes")))
```

And again for the Asian mainland group
```{r, eval=T}
as.bt <- branching.times(extract.clade(vtree, 
                                       getMRCA(vtree, c("dumerilii", "bengalensis"))))
names(as.bt) <- NULL; as.bt <- sort(as.bt, decreasing=T)

vr.bt <- max(nodeHeights(vtree)) - 
         nodeheight(vtree, getParent(vtree, which(vtree$tip.label == "rudicollis")))
vst.bt <- max(nodeHeights(vtree)) - 
         nodeheight(vtree, getParent(vtree, which(vtree$tip.label == "salvator")))
```

Once more for the combined Incumbent vs. Immigrant scenario
```{r}
inc.bt <- c(ov.bt, pa.bt, as.bt)
```


## Making a DAISIE datatable
<!--Now that we have the info, put it into a dataframe for DAISIE.-->
We know the *Odatria/Varanus* radiation is endemic to Australia <!--(minus a couple lineages that have since left - we need to think about how to handle these)-->. *V. keithhornei* is a Cape York endemic species as well, but *V. doreanus* and *V. chlorostigma* exist elsewhere as well (New Guinea). We will characterize these latter two in our data as "Non_endemic_MaxAge", because we know they aren't endemic lineages, but we don't know how old the Australian population is, though it must be younger than the age of their MRCAs.
```{r, eval=T}
DAISIE.oz <- data.frame(Clade_name = c("Odatria_Varanus", 
                                       "keithhornei", 
                                       "doreanus", 
                                       "chlorostigma"),
                        Status = c("Endemic", 
                                   "Endemic", 
                                   "Non_endemic_MaxAge", 
                                   "Non_endemic_MaxAge"),
                        Missing_species = c(0, 0, 0, 0),
                        Branching_times = c(paste(ov.bt, collapse=","), 
                                            vk.bt, vd.bt, vc.bt))
DAISIE.oz[1:4,1:3]
```
```{r}
Australia.endemics <- data.frame(Clade_name = "Odatria_Varanus",
                                 Status = "Endemic",
                                 Missing_species = 0,
                                 Branching_times = c(paste(ov.bt, collapse=",")))
Australia.immigrants <- data.frame(Clade_name = c("keithhornei", "doreanus", "chlorostigma"),
                                   Status = c("Endemic", "Non_endemic_MaxAge", "Non_endemic_MaxAge"),
                                   Missing_species = c(0, 0, 0),
                                   Branching_times = c(vk.bt, vd.bt, vc.bt))
```


View what the distribution of taxa ages and status looks like
```{r}
DAISIE_plot_island(DAISIE.oz, island_age = 35)
```


\  

If you'd like, we can write this dataframe to a file which we can read in later.
```{r, eval=F}
write.table(DAISIE.oz, file = "DAISIE_oz_datatable.txt", 
            sep = "\t", quote = F, row.names = F)
```

\  

Let's do the same for the *Papuan* radiation. The incumbent (endemic) radiation is Hapturosaurus + Euprepiosaurus (mangrove and tree monitors), and then *V. salvadorii*, *V. scalaris*, and *V. panoptes* have emigrated from Australia, and *V.salvator* has emigrated from Indonesia. All three of the four species exist elsewhere and we don't know how long they've been in New Guinea, but the fourth *V. salvadorii* is an endemic lineage.
```{r, eval=T}
DAISIE.pa <- data.frame(Clade_name = c("Mangrove_Tree", 
                                       "salvadorii", 
                                       "scalaris", 
                                       "salvator1",
                                       "panoptes"),
                        Status = c("Endemic", 
                                   "Endemic", 
                                   "Non_endemic_MaxAge", 
                                   "Non_endemic_MaxAge",
                                   "Non_endemic_MaxAge"),
                        Missing_species = c(0, 0, 0, 0, 0),
                        Branching_times = c(paste(pa.bt, collapse=","), 
                                            vsa.bt, vsc.bt, vst.bt, vpa.bt))
DAISIE.pa[1:5,1:3]
```

```{r}
NewGuinea.endemics <- data.frame(Clade_name = "Mangrove_Tree",
                                 Status = "Endemic",
                                 Missing_species = 0,
                                 Branching_times = c(paste(pa.bt, collapse=",")))
NewGuinea.immigrants <- data.frame(Clade_name = c("salvadorii", "scalaris", "salvator1", "panoptes"),
                                   Status = c("Endemic", "Non_endemic_MaxAge", "Non_endemic_MaxAge", "Non_endemic_MaxAge"),
                                   Missing_species = c(0, 0, 0, 0),
                                   Branching_times = c(vsa.bt, vsc.bt, vst.bt, vpa.bt))
```

View what the distribution of taxa ages and status looks like
```{r}
DAISIE_plot_island(DAISIE.pa, island_age = 35)
```

\  

If you'd like, we can write this dataframe to a file which we can read in later.
```{r, eval=F}
write.table(DAISIE.pa, file = "Data/DAISIE_pa_datatable.txt", 
            sep = "\t", quote = F, row.names = F)
```

\  

And one more time for the mainland Asian radiation. The incumbent (endemic) radiation is Empagusia, and *V. rudicollis* and *V. salvator* have arrived from the east. *V. rudicollis* is endemic, and like in the Papuan situation, *V. salvator* is an immigrant and we don't know how long they've been around for.
```{r, eval=T}
DAISIE.as <- data.frame(Clade_name = c("Empagusia", 
                                       "rudicollis", 
                                       "salvator2"),
                        Status = c("Endemic", 
                                   "Endemic", 
                                   "Non_endemic_MaxAge"),
                        Missing_species = c(0, 0, 0),
                        Branching_times = c(paste(as.bt, collapse=","), 
                                            vr.bt, vst.bt))
DAISIE.as[1:3,1:3]
```

```{r}
Asia.endemics <- data.frame(Clade_name = "Empagusia",
                                 Status = "Endemic",
                                 Missing_species = 0,
                                 Branching_times = c(paste(as.bt, collapse=",")))
Asia.immigrants <- data.frame(Clade_name = c("rudicollis", "salvator2"),
                                   Status = c("Endemic", "Non_endemic_MaxAge"),
                                   Missing_species = c(0, 0),
                                   Branching_times = c(vr.bt, vst.bt))
```

View what the distribution of taxa ages and status looks like
```{r}
DAISIE_plot_island(DAISIE.as, island_age = 35)
```

\  

If you'd like, we can write this dataframe to a file which we can read in later.
```{r, eval=F}
write.table(DAISIE.as, file = "DAISIE_as_datatable.txt", 
            sep = "\t", quote = F, row.names = F)
```

\  

Let's do this one last time grouping all the incumbent groups together (Empagusia, Odatria/Varanus, Hapturosaurus/Euprepiosaurus), and all the immigrants together, to see if we can find stronger signal in this combined dataset.
```{r}
DAISIE.combo <- data.frame(Clade_name = c("Odatrai_Varanus",
                                          "keithhornei", 
                                          "doreanus", 
                                          "chlorostigma",
                                          "Mangrove_Tree",
                                          "salvadorii", 
                                          "scalaris", 
                                          "salvator1",
                                          "panoptes",
                                          "Empagusia",
                                          "rudicollis", 
                                          "salvator2"),
                           Status = c("Endemic", 
                                      "Endemic", 
                                      "Non_endemic_MaxAge", 
                                      "Non_endemic_MaxAge",
                                      "Endemic", 
                                      "Endemic", 
                                      "Non_endemic_MaxAge", 
                                      "Non_endemic_MaxAge",
                                      "Non_endemic_MaxAge",
                                      "Endemic", 
                                      "Endemic", 
                                      "Non_endemic_MaxAge"),
                           Missing_species = c(0,0,0,0,0,0,0,0,0,0,0,0),
                           Branching_times = c(paste(ov.bt, collapse=","), vk.bt, vd.bt, vc.bt,
                                               paste(pa.bt, collapse=","), vsa.bt, vsc.bt, vst.bt, vpa.bt,
                                               paste(as.bt, collapse=","), vr.bt, vst.bt))
```

View what the distribution of taxa ages and status looks like
```{r}
DAISIE_plot_island(DAISIE.combo, island_age = 35)
```

---

\  

## Making a DAISIE datalist

Translate your data from a dataframe into a DAISIE datalist.   
This first datalist assumes all taxa belong to a single grouping for the purposes of estimating parameters.  
We'll do this for each of the areas we're investigating. This process also requires that we designate an island age, and the number of lineages that could be in the "source" population (really just number of total number of species - endemics, I think). 
```{r, eval=T}
all.list.oz    <- DAISIE_dataprep(datatable = DAISIE.oz,
                              island_age = 35, M = 30)
#names(all.list.oz) <- rep("Australia", length(all.list.oz))

all.list.pa    <- DAISIE_dataprep(datatable = DAISIE.pa,
                              island_age = 35, M = 50)
#names(all.list.pa) <- rep("NewGuinea", length(all.list.pa))

all.list.as    <- DAISIE_dataprep(datatable = DAISIE.as,
                              island_age = 35, M = 68)
#names(all.list.as) <- rep("Asia", length(all.list.as))

all.list.combo <- list(Australia = all.list.oz,
                       NewGuinea = all.list.pa,
                       Asia = all.list.as)

AUin.list <- DAISIE_dataprep(datatable = Australia.endemics,
                             island_age = 35, M = 30)
AUim.list <- DAISIE_dataprep(datatable = Australia.immigrants,
                             island_age = 35, M = 30)
NGin.list <- DAISIE_dataprep(datatable = NewGuinea.endemics,
                             island_age = 35, M = 50)
NGim.list <- DAISIE_dataprep(datatable = NewGuinea.immigrants,
                             island_age = 35, M = 50)
ASin.list <- DAISIE_dataprep(datatable = Asia.endemics,
                             island_age = 35, M = 68)
ASim.list <- DAISIE_dataprep(datatable = Asia.immigrants,
                             island_age = 35, M = 68)

combined.list <- list(AUin = AUin.list, AUim = AUim.list,
                      NGin = NGin.list, NGim = NGim.list,
                      ASin = ASin.list, ASim = ASim.list)
```

\  

Create a second DAISIE datalist with two groupings (1:Odatria/Varanus, 2:all other immigrant groups; 1:Euprepiosaurus/Hapturosaurus, 2:all immigrants; 1:Empagusia, 2:all immigrants). This will allow us to estimate separate parameter values for the two groups (incumbent vs. immigrant).  
Here we also have to designate the proportion of the source species that the immigrants are (n-immigrants / n-source), which is *prop_type2_pool*.
```{r, eval=T}
split.list.oz <- DAISIE_dataprep(datatable = DAISIE.oz,
                              island_age = 35, M = 30,
                              number_clade_types = 2,
                              list_type2_clades = c("keithhornei", 
                                                    "doreanus", 
                                                    "chlorostigma"),
                              prop_type2_pool = 0.1)

split.list.pa <- DAISIE_dataprep(datatable = DAISIE.pa,
                              island_age = 35, M = 50,
                              number_clade_types = 2,
                              list_type2_clades = c("salvadorii", 
                                                    "scalaris", 
                                                    "salvator1",
                                                    "panoptes"),
                              prop_type2_pool = 0.08)

split.list.as <- DAISIE_dataprep(datatable = DAISIE.as,
                              island_age = 35, M = 68,
                              number_clade_types = 2,
                              list_type2_clades = c("rudicollis", 
                                                    "salvator2"),
                              prop_type2_pool = 0.03)

#split.list.ic <- DAISIE_dataprep(datatable = DAISIE.combo,
#                              island_age = 35, M = 75,
#                              number_clade_types = 2,
#                              list_type2_clades = c("keithhornei", "doreanus", "chlorostigma",
#                                                    "salvadorii", "scalaris", "salvator1", "panoptes",
#                                                    "rudicollis", "salvator2"),
#                              prop_type2_pool = 0.1)

#split.list.ic <- list(Australia = split.list.oz,
#                      NewGuinea = split.list.pa,
#                      Asia = split.list.as)
```

\  

Check out the info that's held in the DAISIE datalist
```{r, eval=T}
# island info
split.list.oz[[1]]

# group 1 info
split.list.oz[[2]][1:2]
```

\pagebreak

# Fitting DAISIE models

Below are examples of how to build and fit a number of models with DAISIE. I've left *eval = F* at the moment because fitting the model a single time is probably insufficient, as maximum likelihood can get trapped in local (sub)optima. We will come back to this in a minute.  

Either way, fitting the models takes quite a bit of time, so my recommendation is to run them externally, then load the model fit objects and summarize/compare/plot them here to keep things clea

--- 

## Equal Rates model with Diversity Dependence
5 parameters (cladogenetic speciation, extinction, carrying capacity, immigration rate, anagenetic speciation) are estimated across ***ALL*** taxa and groups.
```{r, eval=F}
eqr_0 <- DAISIE_ML(datalist = all.list.oz,
                 initparsopt = c(0.2, 0.15, 30, 0.01, 0.1),
                 idparsopt = 1:5,
                 ddmodel = 11,
                 parsfix = NULL, 
                 idparsfix = NULL)
```

--- 

## Equal Rates model without Diversity Dependence
4 parameters (cladogenetic speciation, extinction, immigration rate, anagenetic speciation) are estimated across ***ALL*** taxa and groups. There is no carrying capacity (param 3), so it is fixed at *Infinity*.
```{r, eval=F}
eqr_noDD_0 <- DAISIE_ML(datalist = all.list.oz,
                      initparsopt = c(0.2, 0.15, 0.01, 0.01),
                      idparsopt = c(1,2,4,5),
                      parsfix = Inf,
                      ddmodel = NULL,
                      idparsfix = 3)
```

--- 

## Two Rate Speciation model with Diversity Dependence
4 parameters (extinction, carrying capacity, immigration rate, anagenetic speciation) are estimated across ***ALL*** taxa and groups, but we estimate separate cladogenetic speciation rates for the two groups (*Odatria/Varanus* and immigrants).
```{r, eval=F}
two_lambda <- DAISIE_ML(datalist = split.list.oz,
                        ddmodel = 11,
                        initparsopt = c(1, 0.1, 30, 0.005, 0.1, 
                                        1),
                        idparsopt = c(1,2,3,4,5,6),
                        parsfix = NULL,
                        idparsfix = NULL,
                        idparsnoshift = c(7,8,9,10))
```

--- 

## Two Rate Speciation and Extinction model with Diversity Dependence
3 parameters (carrying capacity, immigration rate, anagenetic speciation) are estimated across ***ALL*** taxa and groups, but we estimate separate cladogenetic speciation and extinction rates for the two groups (*Odatria/Varanus* and immigrants).
```{r, eval=F}
two_lambda_mu <- DAISIE_ML(datalist = split.list.oz,
                           ddmodel = 11,
                           initparsopt = c(1, 0.1, 30, 0.005, 0.1, 
                                           1, 0.1),
                           idparsopt = c(1,2,3,4,5,6,7),
                           parsfix = 0.8,
                           idparsfix = 11,
                           idparsnoshift = c(8,9,10))
```

--- 

## Comparing the fit of DAISIE Models
Compare the fit of the models using the custom *DAISIE.AIC* function.
```{r, eval=T}
source("../Scripts/Calculate_AICs.R")
#DAISIE.AIC(models = c("eqr", "two_lambda"))
#DAISIE.AIC(models = c("eqr", "eqr_noDD"))
```

\pagebreak

# Fitting DAISIE Models in Parallel

Import a custom function to fit DAISIE models multiply from different starting parameter values
```{r, eval=T}
source("../Scripts/search.surface.DAISIE.R")
```

\  

I'll try and quickly explain what you need to provide the function:  
    $\bullet$ **model**---the DAISIE datalist suited for the model you'd like to fit  
    $\bullet$ **n.iter**---the number of model fittings you'd like completed, defaults to 10  
    $\bullet$ **n.proc**---number of processors. this function will fit the model in parallel  
    $\bullet$ **g1.params**---if you'd like to provide starting values for parameters 1:5 (lambda_c, mu, K, gamma, lambda_a) do that here. Otherwise, they'll be drawn randomly from a reasonable (flat) prior  
    $\bullet$ **g1.params**---if you need to estimate parameters for a second group 6:10 (lambda_c, mu, K, gamma, lambda_a) do that here. They'll be drawn randomly from a reasonable (flat) prior. I could design it to work like g1.params, but I just haven't yet  
    $\bullet$ **results**---would you like the function to report just the best fitting run, or all the results from each fit attempt  
    $\bullet$ **start.params**---ignore this, just leave it as NULL  
    $\bullet$ **dd.model**---same as *ddmodel* in DAISIE, 11 for linear density dependence, etc.  
    $\bullet$ **id.parsopt**---same as *idparsopt* in DAISIE, ids for which parameters to optimize  
    $\bullet$ **pars.fix**---same as *parsfix* in DAISIE, which parameters to fix  
    $\bullet$ **id.parsfix**---same as *idparsfix* in DAISIE, ids for the parameters to fix  
    
--- 

## Equal Rates model with Diversity Dependence
5 parameters (cladogenetic speciation, extinction, carrying capacity, immigration rate, anagenetic speciation) are estimated across ***ALL*** taxa and groups.
```{r, eval=F}
eqr <- search.surface.DAISIE(model = all.list.oz, n.iter = 10, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", start.params = NULL,
                             dd.model = 11,
                             id.parsopt = c(1:5),
                             pars.fix = NULL,
                             id.parsfix = NULL,
                             id.parsnoshift = NULL,
                             no.types = 1)
```

If you fit the model with *results = "all"*, then you'll get all the model fittings, plus a designated *best.result* which we can look at below:
```{r, eval=F}
eqr$best.result
```
Otherwise, if you fit *results = "best"*, then you only return the best fitting version.

--- 

Before we get going, let's designate the group we want to work with
```{r, eval=F}
curr.group <- all.list.pa
curr.split <- split.list.pa
ntaxa <- 23
samp.prop <- 0.9
```


## Pure Birth Models
In all Pure Birth models, the extinction rate is 0, so this parameter is not estimated (fixed).

### Pure-Birth **without** Diversity Dependence
This model estimates a single speciation rate shared by the two groups, extinction is set to zero, and there is no diversity dependence (carrying capacity *K* = Infinite).
```{r, eval=F}
eqpb_noDD <- search.surface.DAISIE(model = curr.group, n.iter = 10, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             dd.model = 0,
                             id.parsopt = c(1,4,5),
                             pars.fix = c(0,Inf),
                             id.parsfix = c(2,3),
                             no.types = 1,
                             nsamples = ntaxa)
```

### PB with Speciation Linear to DD
Pure-Birth model with equal speciation rates linear to Diversity Dependence
This model estimates a single speciation rate shared by the two groups, extinction is set to zero, and the speciation rate is linearly dependent on the carrying capacity *K* (linear diversity dependence, ddmodel = 1).
```{r, eval=F}
eqpb_linDD_1 <- search.surface.DAISIE(model = curr.group, n.iter = 10, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             dd.model = 1,
                             id.parsopt = c(1,3,4,5),
                             pars.fix = 0,
                             id.parsfix = 2,
                             no.types = 1,
                             nsamples = ntaxa)
```

### PB with Speciation Exponential to DD
Pure-Birth model with equal speciation rates exponential to Diversity Dependence
This model estimates a single speciation rate shared by the two groups, extinction is set to zero, and the speciation rate is exponentially dependent on the carrying capacity *K* (exponential diversity dependence, ddmodel = 2).
```{r, eval=F}
eqpb_expDD_2 <- search.surface.DAISIE(model = curr.group, n.iter = 10, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             dd.model = 2,
                             id.parsopt = c(1,3,4,5),
                             pars.fix = 0,
                             id.parsfix = 2,
                             no.types = 1,
                             nsamples = ntaxa)
```

### PB with Speciation/Immigration Linear to DD
Pure-Birth model with equal speciation rates and immigration rate linear to Diversity Dependence
This model estimates a single speciation rate shared by the two groups, extinction is set to zero, and the speciation and immigration rates are linearly dependent on the carrying capacity *K* (linear diversity dependence, ddmodel = 11).
```{r, eval=F}
eqpb_linDD_11 <- search.surface.DAISIE(model = curr.group, n.iter = 5, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             dd.model = 11,
                             id.parsopt = c(1,3,4,5),
                             pars.fix = 0,
                             id.parsfix = 2,
                             no.types = 1,
                             nsamples = ntaxa)
```

### PB with Speciation/Immigration Exponential to DD
Pure-Birth model with equal speciation rates and immigration rate exponential to Diversity Dependence
This model estimates a single speciation rate shared by the two groups, extinction is set to zero, and the speciation and immigration rates are exponentially dependent on the carrying capacity *K* (exponential diversity dependence, ddmodel = 21).
```{r, eval=F}
eqpb_expDD_21 <- search.surface.DAISIE(model = curr.group, n.iter = 5, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             dd.model = 21,
                             id.parsopt = c(1,3,4,5),
                             pars.fix = 0,
                             id.parsfix = 2,
                             no.types = 1,
                             nsamples = ntaxa)
```


```{r, eval=F, include=F}
pb_res <- list(eqpb_noDD = eqpb_noDD,
               eqpb_linDD_1 = eqpb_linDD_1,
               eqpb_expDD_2 = eqpb_expDD_2,
               eqpb_linDD_11 = eqpb_linDD_11,
               eqpb_expDD_21 = eqpb_expDD_21)
saveRDS(pb_res, "../Results/DAISIE_Australian_PureBirth_Results.RDS")
```

\pagebreak

## Birth Death Models
In all Birth Death models **both** speciation and extinction are estimated from the data.

### BD without DD
Birth Death model with equal speciation/extinction rates **without** Diversity Dependence
Note: this took ~2 min.
```{r, eval=F}
eqr_noDD <- search.surface.DAISIE(model = curr.group, n.iter = 10, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             #start.params = unlist(eqr_noDD_0[c(1,2,4,5)]),
                             dd.model = 0,
                             id.parsopt = c(1,2,4,5),
                             pars.fix = Inf,
                             id.parsfix = 3,
                             no.types = 1,
                             nsamples = ntaxa)
```

### BDeq with Speciation Linear to DD
Birth Death model with equal speciation/extinction rates, and specation linear to Diversity Dependence
Note: this took ~1 min.
```{r, eval=F}
eqr_linDD_1 <- search.surface.DAISIE(model = curr.group, n.iter = 10, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             dd.model = 1,
                             id.parsopt = c(1:5),
                             pars.fix = NULL,
                             id.parsfix = NULL,
                             no.types = 1,
                             nsamples = ntaxa)
```

### BDeq with Speciation Exponential to DD
Birth Death model with equal speciation/extinction rates, and speciation exponential to Diversity Dependence
Note: this took ~2 min
```{r, eval=F}
eqr_expDD_2 <- search.surface.DAISIE(model = curr.group, n.iter = 5, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             dd.model = 2,
                             id.parsopt = c(1:5),
                             pars.fix = NULL,
                             id.parsfix = NULL,
                             no.types = 1,
                             nsamples = ntaxa)
```

### BDeq with Speciation/Immigration Linear to DD
Birth Death model with equal speciation/extinction rates, and speciation/immigration linear to Diversity Dependence
Note: this took ~35 seconds
```{r, eval=F}
eqr_linDD_11 <- search.surface.DAISIE(model = curr.group, n.iter = 10, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             dd.model = 11,
                             id.parsopt = c(1:5),
                             pars.fix = NULL,
                             id.parsfix = NULL,
                             no.types = 1,
                             nsamples = ntaxa)
```

### BDeq with Speciation/Immigration Exponential to DD
Birth Death model with equal speciation/extinction rates, and speciation/immigration exponential to Diversity Dependence
Note: this took ~2 min
```{r, eval=F}
eqr_expDD_21 <- search.surface.DAISIE(model = curr.group, n.iter = 10, n.proc = 12,
                             g1.params = NULL, g2.params = NULL,
                             results = "all", 
                             start.params = NULL,
                             dd.model = 21,
                             id.parsopt = c(1:5),
                             pars.fix = NULL,
                             id.parsfix = NULL,
                             no.types = 1,
                             nsamples = ntaxa)
```

Save all the results of the Equal Rates models to a file
```{r, eval=F, include=F}
eqr_res <- list(eqr_noDD = eqr_noDD,
                eqr_linDD_1 = eqr_linDD_1,
                eqr_expDD_2 = eqr_expDD_2,
                eqr_linDD_11 = eqr_linDD_11,
                eqr_expDD_21 = eqr_expDD_21)
saveRDS(eqr_res, "../Results/DAISIE_Australian_EqualRatesBD_Results.RDS")
```

\pagebreak

## Split-Rate Birth Death Models
These models follow the birth death models above, but allow different rates (speciation, extinction, or both) between incumbent and immigrant lineages. They require a DAISIE datalist with two groups designed using *DAISIE_dataprep( )* with *number_clade_types* = 2. We'll have 3 sets of these, allowing (1) **speciation** rates *only* to vary between groups, (2) **extinction** rates *only* to vary between groups, and (3) **speciation** and **extinction** rates to *both* vary between groups. We'll include two additional model sets where speciation rate in the immigrant clade is constrained to 0. One thing to remember is that it's necessary to fix the parameter 11 *p_f* in split-rate models. This is the fraction of mainland species which belong to the second subset rate category. For example, in New Guinea, there are 23 species of Varanus, but 19 of them are in the Hapturosaurus/Euprepiosaurus clade. So when analyzing the Australian dataset, the *p_f* of the immigrant taxa (all Hap/Eup) will be 0.82. 

### Immigrant Speciation 0 (ImmSp0)
We can start with a very simplistic split rate model where the speciation rate in the immigrant groups is 0.
```{r, eval=F}
onesp_noDD <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = NULL,
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 0,
                                    id.parsopt = c(1,2,4,5),
                                    pars.fix = c(Inf,0,samp.prop),
                                    id.parsfix = c(3,6,11),
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### Incumbent Speciation Linear to DD, ImmSp0
```{r, eval=F}
onesp_linDD_1 <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = NULL,
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 1,
                                    id.parsopt = c(1,2,3,4,5),
                                    pars.fix = c(0,samp.prop),
                                    id.parsfix = c(6,11),
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

```{r, eval=F}
onesp_fixK_linDD_1 <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = NULL,
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 1,
                                    id.parsopt = c(1,2,4,5),
                                    pars.fix = c(41,0,samp.prop),
                                    id.parsfix = c(3,6,11),
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### Incumbent Speciation Exponential to DD, ImmSp0
```{r, eval=F}
onesp_expDD_2 <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = NULL,
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 2,
                                    id.parsopt = c(1,2,3,4,5),
                                    pars.fix = c(0,samp.prop),
                                    id.parsfix = c(6,11),
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### Incumbent Speciation/Immigration Linear to DD, ImmSp0
```{r, eval=F}
onesp_linDD_11 <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = NULL,
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 11,
                                    id.parsopt = c(1,2,3,4,5),
                                    pars.fix = c(0,samp.prop),
                                    id.parsfix = c(6,11),
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### Incumbent Speciation/Immigration Exponential to DD, ImmSp0
```{r, eval=F}
onesp_expDD_21 <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = NULL,
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 21,
                                    id.parsopt = c(1,2,3,4,5),
                                    pars.fix = c(0,samp.prop),
                                    id.parsfix = c(6,11),
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

\pagebreak

### BD2sp without DD
Birth Death model with **two** speciation rates **without** diversity dependence
4 parameters (extinction, carrying capacity, immigration rate, anagenetic speciation) are estimated across ***ALL*** taxa and groups, but we estimate separate cladogenetic speciation rates for the two groups (*Odatria/Varanus* and immigrants).
Note: this took ~2 min
```{r, eval=F}
twosp_noDD <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 0,
                                    id.parsopt = c(1,2,4,5,6),
                                    pars.fix = c(Inf,samp.prop),
                                    id.parsfix = c(3,11),
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2sp Speciation Linear to DD
BD model with **two** speciation rates linear to Diversity Dependence
Note: this took ~45 sec
```{r, eval=F}
twosp_linDD_1 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 1,
                                    id.parsopt = c(1:6),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2sp Speciation Exponential to DD
BD model with **two** speciation rates exponential to Diversity Dependence
Note: this took ~2 min
```{r, eval=F}
twosp_expDD_2 <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 2,
                                    id.parsopt = c(1:6),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2sp Speciation/Immigration Linear to DD
BD model with **two** speciation/immigration rates linear to Diversity Dependence
Note: this took ~1 min
```{r, eval=F}
twosp_linDD_11 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 11,
                                    id.parsopt = c(1:6),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2sp Speciation/Immigration Exponential to DD
BD model with **two** speciation/immigration rates exponential to Diversity Dependence
Note: this took ~2.5 min
```{r, eval=F}
twosp_expDD_21 <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 21,
                                    id.parsopt = c(1:6),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(7:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```


Save all the results of the Two Sp Rates models to a file
```{r, eval=F, include=F}
twosp_res <- list(twosp_noDD =     twosp_noDD,
                  twosp_linDD_1 =  twosp_linDD_1,
                  twosp_expDD_2 =  twosp_expDD_2,
                  twosp_linDD_11 = twosp_linDD_11,
                  twosp_expDD_21 = twosp_expDD_21)
saveRDS(twosp_res, "../Results/DAISIE_Australian_TwoSpRatesBD_Results.RDS")
```

\pagebreak

### BD2ex without DD
BD model with two extinction rates (equal speciation rates) and no Diversity Dependence
~35 min
```{r, eval=F}
twoex_noDD <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    # becuase the way I've got the 'init.params' bit coded
                                    # to optimize a separate extinction but not speciation
                                    # you still have to designate speciation in the g2.params
                                    # but speciation is listed in the 'id.parsnoshift' set
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 11,
                                    id.parsopt = c(1,2,4,5,7),
                                    pars.fix = c(Inf, samp.prop),
                                    id.parsfix = c(3,11),
                                    id.parsnoshift = c(6,8,9,10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2ex with Speciation Linear to DD
BD model with two extinction rates (equal speciation rates) and speciation rate linear to Diversity Dependence
~ 1 min
```{r, eval=F}
twoex_linDD_1 <- search.surface.DAISIE(model = curr.split, n.iter = 10, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    # becuase the way I've got the 'init.params' bit coded
                                    # to optimize a separate extinction but not speciation
                                    # you still have to designate speciation in the g2.params
                                    # but speciation is listed in the 'id.parsnoshift' set
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 1,
                                    id.parsopt = c(1,2,3,4,5,7),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(6,8,9,10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2ex with Speciation Exponential to DD
BD model with two extinction rates (equal speciation rates) and speciation rate exponential to Diversity Dependence
~1.5  min
```{r, eval=F}
twoex_expDD_2 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    # becuase the way I've got the 'init.params' bit coded
                                    # to optimize a separate extinction but not speciation
                                    # you still have to designate speciation in the g2.params
                                    # but speciation is listed in the 'id.parsnoshift' set
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 2,
                                    id.parsopt = c(1,2,3,4,5,7),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(6,8,9,10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2ex with Speciation/Immigration Linear to DD
BD model with two extinction rates (equal speciation rates) and speciation/immigration rates linear to Diversity Dependence
~ 1min
```{r, eval=F}
twoex_linDD_11 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    # becuase the way I've got the 'init.params' bit coded
                                    # to optimize a separate extinction but not speciation
                                    # you still have to designate speciation in the g2.params
                                    # but speciation is listed in the 'id.parsnoshift' set
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 11,
                                    id.parsopt = c(1,2,3,4,5,7),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(6,8,9,10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2ex with Speciation/Immigration Exponential to DD
BD model with two extinction rates (equal speciation rates) and speciation/immigration rates linear to Diversity Dependence
~ 2 min
```{r, eval=F}
twoex_expDD_21 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    # becuase the way I've got the 'init.params' bit coded
                                    # to optimize a separate extinction but not speciation
                                    # you still have to designate speciation in the g2.params
                                    # but speciation is listed in the 'id.parsnoshift' set
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 21,
                                    id.parsopt = c(1,2,3,4,5,7),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(6,8,9,10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

Save all the results of the Two Ex Rates models to a file
```{r, eval=F, include=F}
twoex_res <- list(twoex_noDD =     twoex_noDD,
                  twoex_linDD_1 =  twoex_linDD_1,
                  twoex_expDD_2 =  twoex_expDD_2,
                  twoex_linDD_11 = twoex_linDD_11,
                  twoex_expDD_21 = twoex_expDD_21)
saveRDS(twoex_res, "../Results/DAISIE_Australian_TwoExRatesBD_Results.RDS")
```

\pagebreak

## Split-Rate Birth Death Models with varying speciation **and** extinction

### ImmSp0, BD2ex, noDD
Two extinction rates, and speciation rate in immigrants is fixed to 0 with no diversity dependence.
~ 45 sec
```{r, eval=F}
twoex_immsp0_noDD <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 0,
                                    id.parsopt = c(1,2,4,5,7),
                                    pars.fix = c(Inf,0,samp.prop),
                                    id.parsfix = c(3,6,11),
                                    id.parsnoshift = c(8:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### ImmSp0, BD2ex, Speciation Linear to DD
~ 40 sec
```{r, eval=F}
twoex_immsp0_linDD_1 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 1,
                                    id.parsopt = c(1,2,3,4,5,7),
                                    pars.fix = c(0,samp.prop),
                                    id.parsfix = c(6,11),
                                    id.parsnoshift = c(8:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### ImmSp0, BD2ex, Speciation Exponential to DD
~ 1.5 sec
```{r, eval=F}
twoex_immsp0_expDD_2 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 2,
                                    id.parsopt = c(1,2,3,4,5,7),
                                    pars.fix = c(0,samp.prop),
                                    id.parsfix = c(6,11),
                                    id.parsnoshift = c(8:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### ImmSp0, BD2ex, Speciation/Immigration Linear to DD
~ 45 sec
```{r, eval=F}
twoex_immsp0_linDD_11 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 11,
                                    id.parsopt = c(1,2,3,4,5,7),
                                    pars.fix = c(0,samp.prop),
                                    id.parsfix = c(6,11),
                                    id.parsnoshift = c(8:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### ImmSp0, BD2ex, Speciation/Immigration Exponential to DD
~  1.7 min
```{r, eval=F}
twoex_immsp0_expDD_21 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 21,
                                    id.parsopt = c(1,2,3,4,5,7),
                                    pars.fix = c(0,samp.prop),
                                    id.parsfix = c(6,11),
                                    id.parsnoshift = c(8:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

\pagebreak

### BD2spex without DD
BD model with two speciation and extinction rates and **without** Diversity Dependence
Note: this took ~ 1 min
```{r, eval=F}
twospex_noDD <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 0,
                                    id.parsopt = c(1,2,4,5,6,7),
                                    pars.fix = c(Inf,samp.prop),
                                    id.parsfix = c(3,11),
                                    id.parsnoshift = c(8:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2spex with Speciation Linear to DD
BD model with two speciation and extinction rates and speciation rates linear to Diversity Dependence
~1.2 min
```{r, eval=F}
twospex_linDD_1 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 1,
                                    id.parsopt = c(1:7),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(8:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2spex with Speciation Exponential to DD
BD model with two speciation and extinction rates and speciation rates exponential to Diversity Dependence
~ 2.4 min
```{r, eval=F}
twospex_expDD_2 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 2,
                                    id.parsopt = c(1:7),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(8:10),
                                    no.types = 2, 
                                    nsamples = ntaxa)
```

### BD2spex with Speciation/Immigration Linear to DD
BD model with two speciation and extinction rates and speciation/immigration rates linear to Diversity Dependence
~ 42 sec
```{r, eval=F}
twospex_linDD_11 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 11,
                                    id.parsopt = c(1:7),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(8:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

### BD2spex with Speciation/Immigration Exponential to DD
BD model with two speciation and extinction rates and speciation/immigration rates exponential to Diversity Dependence
~ 2 min
```{r, eval=F}
twospex_expDD_21 <- search.surface.DAISIE(model = curr.split, n.iter = 5, n.proc = 12,
                                    g1.params = NULL, g2.params = c("lambda_c", "mu"),
                                    results = "all", 
                                    start.params = NULL,
                                    dd.model = 21,
                                    id.parsopt = c(1:7),
                                    pars.fix = samp.prop,
                                    id.parsfix = 11,
                                    id.parsnoshift = c(8:10),
                                    no.types = 2,
                                    nsamples = ntaxa)
```

Save all the results of the Two Sp/Ex Rates models to a file
```{r, eval=F, include=F}
twospex_res <- list(twospex_noDD =     twospex_noDD,
                    twospex_linDD_1 =  twospex_linDD_1,
                    twospex_expDD_2 =  twospex_expDD_2,
                    twospex_linDD_11 = twospex_linDD_11,
                    twospex_expDD_21 = twospex_expDD_21)
saveRDS(twospex_res, "../Results/DAISIE_Australian_TwoSpExRatesBD_Results.RDS")
```


---


Save all the results of the Two Ex Rates models to a file
```{r, eval=F, include=F}
twoex_res <- list(twoex_noDD =     twoex_noDD,
                    twoex_linDD_1 =  twoex_linDD_1,
                    twoex_expDD_2 =  twoex_expDD_2,
                    twoex_linDD_11 = twoex_linDD_11,
                    twoex_expDD_21 = twoex_expDD_21)
saveRDS(twoex_res, "../Results/DAISIE_Australian_TwoExRatesBD_Results.RDS")
```

\pagebreak


Make sure you've got all the models saved somewhere safe
```{r, eval=F}
#all.models <- list(pb_res, eqr_res, twosp_res, twospex_res, twoex_res)
#saveRDS(all.models, "../Results/DAISIE_Australian_AllModel_Results.RDS")

# MAKE SURE THAT YOU CHANGE THE NAME OF THE OUTPUT FILE!
save(eqpb_noDD, eqpb_linDD_1, eqpb_linDD_11, eqpb_expDD_2, eqpb_expDD_21,
     eqr_noDD, eqr_linDD_1, eqr_linDD_11,eqr_expDD_2, eqr_expDD_21, 
     onesp_noDD, onesp_linDD_1, onesp_linDD_11, onesp_expDD_2, onesp_expDD_21,
     twosp_noDD, twosp_linDD_1, twosp_linDD_11, twosp_expDD_2, twosp_expDD_21,
     twoex_noDD, twoex_linDD_1, twoex_linDD_11, twoex_expDD_2, twoex_expDD_21,
     twoex_immsp0_noDD, twoex_immsp0_linDD_1, twoex_immsp0_linDD_11,
     twoex_immsp0_expDD_2, twoex_immsp0_expDD_21,
     twospex_noDD, twospex_linDD_1, twospex_linDD_11, 
     twospex_expDD_2, twospex_expDD_21, 
     file="../Results/DAISIE_NewGuinea_AllModel_Results.RData")
```
---

\pagebreak

# Comparing Empirical Model Fits

Start by designating which models we want to compare.
```{r}
all.DAISIE.models <- c("eqpb_noDD", "eqpb_linDD_1", "eqpb_linDD_11",
                       "eqpb_expDD_2", "eqpb_expDD_21",
                       "eqr_noDD", "eqr_linDD_1", "eqr_linDD_11",
                       "eqr_expDD_2", "eqr_expDD_21",
                       "onesp_noDD", "onesp_linDD_1", "onesp_linDD_11",
                       "onesp_expDD_2", "onesp_expDD_21",
                       "twosp_noDD", "twosp_linDD_1", "twosp_linDD_11",
                       "twosp_expDD_2", "twosp_expDD_21",
                       "twospex_noDD", "twospex_linDD_1", "twospex_linDD_11",
                       "twospex_expDD_2", "twospex_expDD_21",
                       "twoex_noDD", "twoex_linDD_1", "twoex_linDD_11",
                       "twoex_expDD_2", "twoex_expDD_21",
                       "twoex_immsp0_noDD", "twoex_immsp0_linDD_1",
                       "twoex_immsp0_linDD_11", "twoex_immsp0_expDD_2",
                       "twoex_immsp0_expDD_21")
```


## Model Fits with AIC

Use the *DAISIE.AIC* function to compare the model fits with AIC  
This will take the model names as input, use the likelihood to calculate AIC and corrected AIC (AICc), and tell us the contribution of each model to the total model weights (AICcwt).
```{r, eval = F, include=T}
all.AIC <- DAISIE.AIC(models = all.DAISIE.models, 
                      method="best")
```

\  
\  

## Model Fits with BIC

Use the *DAISIE.BIC* function to compare the model fits with BIC    
This will take the model as input, use the likelihood to calculate BIC, and tell us the contribution of each model to the total model weights (BICwt)
```{r, eval=F, include=T}
all.BIC <- DAISIE.BIC(models = all.DAISIE.models, 
                      method="best")
```

\pagebreak

## **Australia** Results
Let's check model comparison using AIC first
```{r, eval=T, echo=F}
load("../Results/DAISIE_Australian_AllModel_Results.RData")
```

```{r, eval = T, echo=T}
all.AIC <- DAISIE.AIC(models = all.DAISIE.models, 
                      method="best")
```
```{r, eval=T, echo=F}
all.AIC$results[,c(1,2,4,5,6)]
all.AIC$best.model
all.AIC$parameter.estimates
```

\  

Then with BIC next
```{r, eval=T, echo=T}
all.BIC <- DAISIE.BIC(models = all.DAISIE.models, 
                      method="best")
```
```{r, eval=T, echo=F}
all.BIC
```

\pagebreak

```{r, eval=T, echo=F}
load("../Results/DAISIE_NewGuinea_AllModel_Results.RData")
```

## Results for the **New Guinea** Monitors
AIC first
```{r, eval = T, echo=T}
all.AIC <- DAISIE.AIC(models = all.DAISIE.models, 
                      method="best")
```
```{r, eval=T, echo=F}
all.AIC$results[,c(1,2,4,5,6)]
all.AIC$best.model
all.AIC$parameter.estimates
```

\  

BIC next
```{r, eval=T, echo=T}
all.BIC <- DAISIE.BIC(models = all.DAISIE.models, 
                      method="best")
```
```{r, eval=T, echo=F}
all.BIC
```

---

\pagebreak

<!--Let's try modelling the these rates across the combined datasets-->
```{r, eval=F, include=F}
eqr_noDD <- DAISIE_ML(
            datalist = all.list.combo,
            datatype = 'multiple',
            initparsopt = c(0.1, 0.01, 0.05, 0.1),
            idparsmat = rbind(1:5, 1:5, 1:5),
            idparsopt = c(1,2,4,5),
            parsfix = Inf,
            idparsfix = 3
)
```
```{r, eval=F, include=F}
eqr_noDD <- DAISIE_ML(
            datalist = all.list.combo,
            datatype = 'multiple',
            initparsopt = c(0.1, 0.01, 0.05, 0.1),
            idparsmat = rbind(1:5, 1:5, 1:5),
            idparsopt = c(1,2,4,5),
            parsfix = Inf,
            idparsfix = 3
)
```


```{r, eval=F, include=F}
split.ii_noDD <- DAISIE_ML(
                 datalist = combined.list,
                 datatype = 'multiple',
                 initparsopt = c(0.1, 0.01, 0.05, 0.1, 
                                 0.1, 0.01, 0.05, 0.1,
                                 0.),
                 idparsmat = rbind(1:5, 6:10, 1:5, 6:10),
                 idparsopt = c(1,2,4,5),
                 parsfix = Inf,
                 idparsfix = 3
)
```



\newpage
\blandscape

# DAISIE Models Explained

## Model Categories

Q   Regimes   Model Type               Model Name                   Shared                                             Varied               
--  --------  ----------------        -----------------            -----------------                                  ----------           
1   1         Pure Birth              Pure Birth---*eqpb*          $\lambda$^c^, $\gamma$, $\lambda$^a^, *K*          ---                   
2   1         Birth Death             Equal Rates---*eqr*          $\lambda$^c^, $\mu$, $\gamma$, $\lambda$^a^, *K*   --- 
3   2         Split Rate Birth Death  Split $\lambda$---*onesp*    $\mu$, $\gamma$, $\lambda$^a^, *K*                 $\lambda$^c^ ($\lambda$^c^~2~=0)
4   2         Split Rate Birth Death  Split $\lambda$---*twosp*    $\mu$, $\gamma$, $\lambda$^a^, *K*                 $\lambda$^c^            
5   2         Split Rate Birth Death  Split $\mu$---*twoex*        $\lambda$^c^, $\gamma$, $\lambda$^a^, *K*          $\mu$    
6   2         Split Rate Birth Death  Split $\mu$---*twoex_immsp0* $\gamma$, $\lambda$^a^, *K*                        $\lambda$^c^ ($\lambda$^c^~2~=0), $\mu$
7   2         Split Rate Birth Death  Split Sp/Ex---*twospex*      $\gamma$, $\lambda$^a^, *K*                        $\lambda$^c^, $\mu$     


## Model Specifications

Model#   Q   DAISIE Model            Model Name              DD     #Par.  $\lambda$-DD Corr.  $\gamma$-DD Corr.   Fixed Parameters            Free Parameters
------   --  ------------            --------------          ---    -----  ------------------  ----------------    --------------------        ----------------------
M1       1   *eqpb_noDD*             Pure Birth              0      3      ---                 ---                 *K*=Inf                     $\lambda$^c^, $\gamma$, $\lambda$^a^
M2       1   *eqpb_linDD_1*          Pure Birth              1      4      linear              ---                                             $\lambda$^c^, $\gamma$, $\lambda$^a^, *K*
M3       1   *eqpb_linDD_11*         Pure Birth              11     4      linear              linear                                          $\lambda$^c^, $\gamma$, $\lambda$^a^, *K*
M4       1   *eqpb_expDD_2*          Pure Birth              2      4      exponential         ---                                             $\lambda$^c^, $\gamma$, $\lambda$^a^, *K*
M5       1   *eqpb_expDD_21*         Pure Birth              21     4      exponential         exponential                                     $\lambda$^c^, $\gamma$, $\lambda$^a^, *K*
M6       2   *eqr_noDD*              Equal Rates             0      4      ---                 ---                 *K*=Inf                     $\lambda$^c^, $\mu$, $\gamma$, $\lambda$^a^
M7       2   *eqr_linDD_1*           Equal Rates             1      5      linear              ---                                             $\lambda$^c^, $\mu$, $\gamma$, $\lambda$^a^, *K*
M8       2   *eqr_linDD_11*          Equal Rates             11     5      linear              linear                                          $\lambda$^c^, $\mu$, $\gamma$, $\lambda$^a^, *K*
M9       2   *eqr_expDD_2*           Equal Rates             2      5      exponential         ---                                             $\lambda$^c^, $\mu$, $\gamma$, $\lambda$^a^, *K*
M10      2   *eqr_expDD_21*          Equal Rates             21     5      exponential         exponential                                     $\lambda$^c^, $\mu$, $\gamma$, $\lambda$^a^, *K*
M11      3   *onesp_noDD*            Split $\lambda$         0      4      ---                 ---                 *K*=Inf, $\lambda$~c2~=0    $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^
M12      3   *onesp_linDD_1*         Split $\lambda$         1      5      linear              ---                 $\lambda$~c2~=0             $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^, *K*
M13      3   *onesp_linDD_11*        Split $\lambda$         11     5      linear              linear              $\lambda$~c2~=0             $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^, *K*
M14      3   *onesp_expDD_2*         Split $\lambda$         2      5      exponential         ---                 $\lambda$~c2~=0             $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^, *K*
M15      3   *onesp_expDD_21*        Split $\lambda$         21     5      exponential         exponential         $\lambda$~c2~=0             $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^, *K*
M16      4   *twosp_noDD*            Split $\lambda$         0      5      ---                 ---                 *K*=Inf                     $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^, $\lambda$^c^~2~
M17      4   *twosp_linDD_1*         Split $\lambda$         1      6      linear              ---                                             $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^, *K*, $\lambda$^c^~2~
M18      4   *twosp_linDD_11*        Split $\lambda$         11     6      linear              linear                                          $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^, *K*, $\lambda$^c^~2~
M19      4   *twosp_expDD_2*         Split $\lambda$         2      6      exponential         ---                                             $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^, *K*, $\lambda$^c^~2~
M20      4   *twosp_expDD_21*        Split $\lambda$         21     6      exponential         exponential                                     $\lambda$^c^~1~, $\mu$, $\gamma$, $\lambda$^a^, *K*, $\lambda$^c^~2~
M21      5   *twoex_noDD*            Split $\mu$             0      5      ---                 ---                 *K*=Inf                     $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, $\mu$~2~
M22      5   *twoex_linDD_1*         Split $\mu$             1      6      linear              ---                                             $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\mu$~2~
M23      5   *twoex_linDD_11*        Split $\mu$             11     6      linear              linear                                          $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\mu$~2~
M24      5   *twoex_expDD_2*         Split $\mu$             2      6      exponential         ---                                             $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\mu$~2~
M25      5   *twoex_expDD_21*        Split $\mu$             21     6      exponential         exponential                                     $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\mu$~2~
M26      6   *twoex_immsp0_noDD*     Split $\lambda$, $\mu$  0      5      ---                 ---                 *K*=Inf, $\lambda$~c2~=0    $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, $\mu$~2~
M27      6   *twoex_immsp0_linDD_1*  Split $\lambda$, $\mu$  1      6      linear              ---                 $\lambda$~c2~=0             $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\mu$~2~
M28      6   *twoex_immsp0_linDD_11* Split $\lambda$, $\mu$  11     6      linear              linear              $\lambda$~c2~=0             $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\mu$~2~
M29      6   *twoex_immsp0_expDD_2*  Split $\lambda$, $\mu$  2      6      exponential         ---                 $\lambda$~c2~=0             $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\mu$~2~
M30      6   *twoex_immsp0_linDD_21* Split $\lambda$, $\mu$  1      6      exponential         exponential         $\lambda$~c2~=0             $\lambda$^c^, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\mu$~2~
M31      7   *twospex_noDD*          Split $\lambda$, $\mu$  0      6      ---                 ---                 *K*=Inf                     $\lambda$^c^~1~, $\mu$~1~, $\gamma$, $\lambda$^a^, $\lambda$^c^~2~, $\mu$~2~
M32      7   *twospex_linDD_1*       Split $\lambda$, $\mu$  1      7      linear              ---                                             $\lambda$^c^~1~, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\lambda$^c^~2~, $\mu$~2~
M33      7   *twospex_linDD_11*      Split $\lambda$, $\mu$  11     7      linear              linear                                          $\lambda$~c1~, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\lambda$^c^~2~, $\mu$~2~
M34      7   *twospex_expDD_2*       Split $\lambda$, $\mu$  2      7      exponential         ---                                             $\lambda$~c1~, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\lambda$^c^~2~, $\mu$~2~
M35      7   *twospex_expDD_21*      Split $\lambda$, $\mu$  21     7      exponential         exponential                                     $\lambda$~c1~, $\mu$~1~, $\gamma$, $\lambda$^a^, *K*, $\lambda$^c^~2~, $\mu$~2~


\elandscape

--- 

\pagebreak

# Diversification Models with ClaDS

Direct your working directory to the folder that holds the ClaDS scripts
```{r setup2, eval=F, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/Google.Drive/R.Analyses/ClaDS/")
```

Load RPANDA
```{r, eval=T, message=F, warning=F}
library(RPANDA)
```

\  

In case you're jumping straight to here, start by reading in the tree
```{r, eval=T, fig.height=10}
vtree <- read.tree("UsedTree.tre")
```

\  

ClaDS (within RPANDA) has a couple of models we can fit  
  $\bullet$ **ClaDS0**---assumes pure birth process when estimating speciation rates (extinction = 0)  
  $\bullet$ **ClaDS1**---assumes a constant extinction rate, and estimates variable speciation rates  
  $\bullet$ **ClaDS2**---estimates variable speciation and extinction rates, but holds turnover rates constant  
  
\  

Fitting these models takes a lot of time, and generations, so make sure to set the iterations relatively high. You can extend runs if you haven't reached convergence though.

--- 

## ClaDS0---Pure Birth Model
```{r, eval=F}
clads0.varanus <- fit_ClaDS0(tree = vtree,
                           sample_fraction = (76/83),
                           iterations = 10000,
                           thin = 100,
                           pamhLocalName = "local",
                           #it_save = 10000,
                           name = "Monitor_ClaDS0",
                           nCPU = 3)
```

We can check the chains for convergence
```{r, eval=F}
plot_ClaDS_chains(clads2.varanus)
```

Then we can extract the per branch rates, and plot them along the tree
```{r, eval=F}
maps <- getMAPS_ClaDS(clads2.varanus, thin = 10)
plot_ClaDS_phylo(vtree, maps[-(1:4)])
```

--- 

## ClaDS2---Constant Turnover Model

```{r, eval=F}
clads2.varanus <- fit_ClaDS(tree = vtree,
                           sample_fraction = (76/83),
                           iterations = 10000,
                           thin = 100,
                           #it_save = 10000,
                           file_name = "Monitor_ClaDS2",
                           model_id = "ClaDS2",
                           nCPU = 3)
```

We can check the chains for convergence
```{r, eval=F}
plot_ClaDS_chains(clads2.varanus)
```

Then we can extract the per branch rates and plot them along the tree
```{r, eval=F}
maps <- getMAPS_ClaDS(clads2.varanus, thin = 10)
plot_ClaDS_phylo(vtree, maps[-(1:4)])
```

---

Here's an example file I didn't run long enough:
```{r, eval=T}
c2_run1 <- readRDS("~/Documents/GitHub/MonitorDiversification/Data/ClaDS2_Varanus.RDS")
```

```{r, eval=F}
plot_ClaDS_chains(c2_run1)
```

Let's look at the plot anyway, then make sure to run for more generations
```{r, eval=T}
maps <- getMAPS_ClaDS(c2_run1, thin = 10)
plot_ClaDS_phylo(vtree, maps[-(1:4)])
```

\  

The above example was run for 10,000 generations, thin = 100. 
Run it longer by just providing the previous file
```{r, eval=F}
c2_run1_cont <- fit_ClaDS(tree = vtree,
                           sample_fraction = (76/83),
                           iterations = 10000,
                           model_id = "ClaDS2",
                           nCPU = 3,
                          sampler = c2_run1)
```

\  

We can read in a longer pre-cooked file. This one has been run for 200K generations and 3 chains.
```{r, evail = T}
load("~/Documents/GitHub/MonitorDiversification/Data/C2_Varanus_200K.RData")
c2.varanus <- mcmcSampler
maps <- getMAPS_ClaDS(c2.varanus, thin = 10)
plot_ClaDS_phylo(vtree, maps[-(1:4)])
```

\  

We might want to compare the speciation rates across different subgenera, or clades.  
Start by making a function to extract the edge (branch) specific rates from the ClaDS object.
```{r, eval=T}
extract.ClaDS.rates <- function(rates, min.branch, max.branch){
            clade.rates <- rates[(min.branch + 4):(max.branch + 4)]
            return(clade.rates)}
```
This function is just really simple, and excludes the first four parameters (sigma, alpha, epsilon, l_0), when isolating the rates of interest.

\  

Now pick a handful of clades to compare:
```{r, eval=T}
odatria.rates <- extract.ClaDS.rates(maps, 7, 49)
hapturo.rates <- extract.ClaDS.rates(maps, 69, 83)
euprep.rates  <- extract.ClaDS.rates(maps, 83, 101)
varanus.rates <- extract.ClaDS.rates(maps, 50, 66)
sotero.rates  <- extract.ClaDS.rates(maps, 119, 137)
empagus.rates <- extract.ClaDS.rates(maps, 112, 118)
african.rates <- extract.ClaDS.rates(maps, 139, 149)
```

Put the rates together into a single data frame
```{r, eval=T}
all.rates <- data.frame(sp_rates = odatria.rates, group = "Odatria")
all.rates <- add_row(all.rates, 
                     sp_rates = hapturo.rates, 
                     group = "Hapturosaurus")
all.rates <- add_row(all.rates, 
                     sp_rates = euprep.rates , 
                     group = "Euprepiosaurus")
all.rates <- add_row(all.rates, 
                     sp_rates = varanus.rates, 
                     group = "Varanus_Papusaurus")
all.rates <- add_row(all.rates, 
                     sp_rates = sotero.rates , 
                     group = "Soterosaurus")
all.rates <- add_row(all.rates, 
                     sp_rates = empagus.rates, 
                     group = "Empagusia")
all.rates <- add_row(all.rates, 
                     sp_rates = african.rates, 
                     group = "Polydaedalus_Psammosaurus")
```

And plot these to get an idea of variation across the groups:
```{r, eval=T, message=F, warning=F}
library(ggridges)
ggplot(all.rates, aes(x=sp_rates, y=group, 
                       point_color=sp_rates)) +
  geom_density_ridges(jittered_points = T, scale = 2,
                      point_shape = "|", alpha = 0.75,
                      point_size = 2, size = 0.5,
                      position = position_points_jitter(height=0),
                      aes(fill = group), show.legend =F) +
  scale_fill_brewer(palette = "RdYlBu") +
  theme_bw()
```